---
title: "Data Wrangling with dplyr"
author: "Briana Mittleman"
date: "1/13/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Subset rows and filter columns  

What should we do first when we open a new markdown?  


Load libraries: 
```{r}
library(dplyr)
library(tidyverse)
```

Load data:  

```{r}
counts_raw <- read_delim("data/counts-raw.txt.gz", delim="\t")
```

Goals:  

subset rows of a dataframe and select columns of a dataframe  


```{r}
research <- filter(counts_raw, articleType=="Research Article")
```

Filter to research articles from 2006  

```{r}
research_2006 <- filter(research, year==2006)
nrow(research_2006)
```

Filter to articles with a tweet  

```{r}
research_2006_tweet <- filter(research_2006, backtweetsCount>0)
nrow(research_2006_tweet)
```

Research articles from 2006 with a facebook comment  

```{r}
research_2006_fb <- filter(research, year==2006, facebookCommentCount>0)
nrow(research_2006_fb)
```

We can use logical operators:  

AND &
OR |  

Look for research articles in 2006 with a tweet or facebook comment  

```{r}
research_2006_fb_tweet <- filter(research, year==2006 , facebookCommentCount >0 | backtweetsCount >0)  
nrow(research_2006_fb_tweet)
```

Article in 2006 with social media coverage and is about infectious diseases  

```{r}
research_2006_fb_tweet_disease <- filter(research, year==2006,  facebookCommentCount >0 | backtweetsCount>0, grepl("Infectious Diseases", plosSubjectTags))

nrow(research_2006_fb_tweet_disease)
```
Subset columns of a dataframe with function subset  

```{r}
colnames(research)
```
Subset the columns with the article information using the dplyr function select()  

```{r}
article_info <- select(research, doi, pubDate, journal, title, articleType, authorsCount)
colnames(article_info)

```
New dataframe with columns doi:authorsCount  

```{r}
article_info<- select(research, doi:authorsCount)
colnames(article_info)
```
Create a metrics dataframe  

```{r}
metrics <- select(research, contains("Count"))
colnames(metrics)
```
Exclude a column after contains  

```{r}
metrics <- select(research, contains("Count"), -authorsCount)
colnames(metrics)
```
```{r}
metrics <- select(research, contains("Count"), -authorsCount, f1000Factor, wikipediaCites)
colnames(metrics)
```
Journal column  

```{r}
head(select(research, journal))
```
```{r}
head(select(research,3))
```
Slice: Slicing only the rows we want by numbers 

Rows 1:3 article info dataframe  
```{r}
slice(article_info,1:3)
```
Challenge 1:  

```{r}
challenge1=filter(research, year==2009, facebookCommentCount >0 | backtweetsCount >0) 
nrow(challenge1)
```

```{r}
challenge2= filter(research, year==2009, mendeleyReadersCount >0)
nrow(challenge2)
```

```{r}
challenge3= filter(research, pdfDownloadsCount >1000, mendeleyReadersCount > 15, year <2009, wosCountThru2011< 10)
nrow(challenge3)
```

```{r}
str(challenge3)
``` 

#Chaining commands with dplyr  

Remeber in unix we pipe with the | sign. In dplyr we use %>%  

%>% %>% %>% %>% 

Facebook articles from 2006  

```{r}
facebook_2006 <- research %>% filter(year==2006) %>% select(contains("facebook"))

head(facebook_2006)
```

```{r}
research_2006 <- filter(research, year==2006)
facebook_2006 <- select(research_2006, contains("facebook"))
```


Pipe with base R functions as well  

```{r}
research %>% filter(year==2006) %>% select(contains("facebook")) %>% head 
```

Arrange function: Use to sort rows in a dataframe  

```{r}
research %>%
  arrange(authorsCount, wosCountThru2011) %>% 
  select(authorsCount, wosCountThru2011) %>% 
  slice(1:10)
```


Flip arrangement of the sort  

```{r}
research %>% 
  arrange(desc(authorsCount), desc(wosCountThru2011)) %>% 
  select(authorsCount, wosCountThru2011) %>% 
  slice(1:10)

```

```{r}
research %>%
  arrange(desc(wosCountThru2011)) %>% 
  select(title) %>% 
  slice(1:3)
  
```

```{r}
research %>% 
  arrange(desc(authorsCount)) %>% 
  select(authorsCount, title, journal, plosSubjectTags) %>% 
  slice(1:3)
```

#Summarize data with dplyr  

Goals: 

* new columns with mutate  
* summarize with summarize   
* count number of observations n()  
* group data by variable (group_by)  


```{r}
research <- research %>% mutate(weeksSincePublished= daysSincePublished/7)

research %>% select(weeksSincePublished) %>% head()
```

```{r}
research <- research %>% mutate( weeksSincePublished= daysSincePublished/7,
                                 yearsSincePublished=weeksSincePublished/52)

select(research, contains("Since")) %>% slice(1:5)
```

Summarize the data: 

```{r}
summarise(research, plos_mean=mean(plosCommentCount))
```

```{r}
plot_stats<- research %>% summarise(plos_mean=mean(plosCommentCount), 
                                    plos_sd=sd(plosCommentCount))

plot_stats
```


```{r}
research %>% filter(journal=="pone", year==2007) %>% 
  summarise(plos_mean=mean(plosCommentCount),
            plos_sd=sd(plosCommentCount))


```

Add number of observations  

```{r}
research %>% filter(journal=="pone", year==2007) %>% 
  summarise(plos_mean=mean(plosCommentCount),
            plos_sd=sd(plosCommentCount),
            num=n())
```

Use group by to get stats by a grouping of the data  
BaseR
```{r}
mean_metric_per_var <- function(metric, variable){
  result=integer(length(variable))
  for (v in varaible){
    result[v] <- means(metric[variable==v])
  }
return(result)
}
```
dplyr
```{r}
research %>% group_by(journal) %>% 
  summarise(tweets_mean=mean(backtweetsCount))
```

Group by 2 variables  

```{r}
research %>% group_by(journal, year) %>% 
  summarise(tweets_mean=mean(backtweetsCount))
```

```{r}
tweets_per_journal <- research %>%
  group_by(journal) %>% 
  summarise(mean_tw=mean(backtweetsCount),
            sd_tw=sd(backtweetsCount),
            num=n(),
            SEM=sd_tw/sqrt(num)) %>% 
  select(-sd_tw)

tweets_per_journal

```
Change









