---
title: "Plotting with ggplot2"
author: "Briana Mittleman"
date: "1/13/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages 
```{r}
library(tidyverse)
```

Load data:  


```{r}
counts_raw <- read_delim("data/counts-raw.txt.gz", delim="\t")
research <- filter(counts_raw, articleType=="Research Article")
```


#Mapping data to plot aesthetics 

Goals:  

* Map columns of a data frame to plot aesthetics using aes  
* Add layers to plot  
 
 
```{r}
p <- ggplot(data=research, mapping=aes(x=pdfDownloadsCount,
                                       y= wosCountThru2011))
```

```{r}
p
```
 
```{r}
p + geom_point()
```
 
```{r}
p <- p + geom_point()
```
 
```{r}
p <- ggplot(research, aes(x=pdfDownloadsCount,
                          y=wosCountThru2011)) +
  geom_point()

p
```


Change aes of geometry  

```{r}
p <- ggplot(research, aes(x=pdfDownloadsCount,
                          y=wosCountThru2011)) +
  geom_point(aes(color=journal))

p
```

Change size of points by variable  

```{r}
p <- ggplot(research, aes(x=pdfDownloadsCount,
                          y=wosCountThru2011)) +
  geom_point(aes(size=authorsCount))
p
```
Change alpha by variable  

```{r}
p <- ggplot(research, aes(x=pdfDownloadsCount,
                          y=wosCountThru2011))+
  geom_point(aes(alpha=daysSincePublished))
p
```
Change color of all points, not based on data condition  

```{r}
p <- ggplot(research, aes(x=pdfDownloadsCount,
                          y=wosCountThru2011)) +
  geom_point(color="red", aes(size=authorsCount))

p
```


Add another geometry layer:  

Smoothed line 

```{r}
p <- ggplot(research, aes(x=pdfDownloadsCount,
                          y=wosCountThru2011)) +
  geom_point(aes(color=journal)) +
  geom_smooth()

p
```


Color both layers by journal  

```{r}
p <- ggplot(research, aes(x=pdfDownloadsCount,
                          y=wosCountThru2011,
                          color=journal))+
  geom_point()+
  geom_smooth()

p
```

```{r}
p_chal <- ggplot(research, aes(x=daysSincePublished,
                               y=wosCountThru2011)) +
  geom_point(alpha=.5, aes(color=journal)) +
  geom_smooth(color="red")


p_chal
```

#Control plot scales  

* scale how the data mapps to the aes  

* control axis  

* control color scale  

```{r}
p <- ggplot(research, aes(x=pdfDownloadsCount,
                          y=wosCountThru2011)) +
  geom_point(aes(color=journal)) +
  geom_smooth()

p
```

```{r}
p + scale_x_log10() + scale_y_log10()
```

```{r}
p <- ggplot(research, aes(x=log10(pdfDownloadsCount + 1 ),
                          y=log10(wosCountThru2011 + 1 ))) +
  geom_point(aes(color=journal)) +
  geom_smooth()

p
```

```{r}
p <- ggplot(research, aes(x=log10(pdfDownloadsCount + 1 ),
                          y=log10(wosCountThru2011 + 1 ))) +
  geom_point(aes(color=journal)) +
  geom_smooth() +
  scale_x_continuous(breaks = c(1,3), labels = c(10,1000)) +
  scale_y_continuous(breaks=c(1,3), labels=c(10,1000))

p
```

Change the color scale  

```{r}
p + scale_color_grey()
```

```{r}
unique(research$journal)
```


```{r}
p + scale_color_manual(values=c("red", "yellow", "orange", "purple", "blue", "orange", "pink"))
```

```{r}
library("RColorBrewer")
display.brewer.all(type="qual")
```

```{r}
p + scale_color_brewer(palette="Dark2")
```

```{r}
p_chal<-ggplot(research, aes(x=sqrt(pdfDownloadsCount), y=sqrt(wosCountThru2011))) +
  geom_point(aes(color=journal))+
  geom_smooth() +
  scale_color_brewer(palette = "Accent") +
  labs(x="Square Root of Downloads", y="Square root of Citation count thru 2011", title="Meaningful title")

p_chal
```










































 